---
title: "`r params$report_title`"
subtitle: "`r params$subtitle`"
author: "`r paste('Prepared by', paste(c(params$author, params$organization)[nzchar(c(params$author, params$organization))], collapse = ' · '))`"
date: "`r format(as.Date(params$report_date), '%B %d, %Y')`"
params:
  report_title: "Synthetic Health Data Summary"
  subtitle: ""
  memo: "NACRS-DEMO"
  author: "Miss V"
  organization: "The V Lab"
  client: "Dr. ABC"
  client_organization: "ABC Department"
  report_date: "2025-01-01"
  version: "1.0"
  status: "Public example"
  confidentiality: "Synthetic data only"
  note: "Demonstration report using synthetic data only."
  data_dir: "code_examples/example_data"
  data_file: "synthetic_nacrs.xlsx"
  include_summary: true
  include_data_overview: true
  include_appendix: false
  show_code: false
output:
  html_document:
    number_sections: true
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: flatly
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = isTRUE(params$show_code),
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 8,
  fig.height = 5,
  out.width = "100%"
)

# Small helpers shared by the whole report.
is_filled <- function(x) {
  length(x) == 1L && !is.na(x) && nzchar(trimws(as.character(x)))
}

as_flag <- function(x) {
  isTRUE(x) || (is.character(x) && tolower(x) %in% c("true", "yes", "1"))
}

report_datasets <- list()
data_sources <- character()
data_aliases <- character()
data_notes <- character()

find_project_dir <- function(start_dir) {
  current_dir <- normalizePath(start_dir, winslash = "/", mustWork = TRUE)

  repeat {
    r_projects <- list.files(
      current_dir,
      pattern = "\\.[Rr]proj$",
      full.names = TRUE
    )

    if (length(r_projects) > 0L || dir.exists(file.path(current_dir, ".git"))) {
      return(current_dir)
    }

    parent_dir <- dirname(current_dir)
    if (identical(parent_dir, current_dir)) {
      stop("Could not locate the project root.", call. = FALSE)
    }

    current_dir <- parent_dir
  }
}

input_file <- knitr::current_input(dir = TRUE)
input_dir <- if (nzchar(input_file)) {
  dirname(normalizePath(input_file, winslash = "/", mustWork = TRUE))
} else {
  getwd()
}
project_dir <- find_project_dir(input_dir)
data_dir_display <- params$data_dir
source(
  file.path(project_dir, "code_examples", "example_helpers.R"),
  local = TRUE
)

if (!requireNamespace("readxl", quietly = TRUE)) {
  data_notes <- paste(
    "Loading the example workbook requires the 'readxl' package.",
    "Run 'Rscript code_examples/setup_dependencies.R' first."
  )
} else if (is_filled(params$data_dir)) {
  expanded_data_dir <- path.expand(params$data_dir)
  is_absolute_data_dir <- grepl(
    "^(?:/|[A-Za-z]:[/\\\\]|\\\\\\\\)",
    expanded_data_dir
  )
  resolved_data_dir <- if (is_absolute_data_dir) {
    expanded_data_dir
  } else {
    file.path(project_dir, expanded_data_dir)
  }
  data_dir <- normalizePath(
    resolved_data_dir,
    winslash = "/",
    mustWork = FALSE
  )
  project_prefix <- paste0(project_dir, "/")
  data_dir_display <- if (startsWith(data_dir, project_prefix)) {
    substring(data_dir, nchar(project_prefix) + 1L)
  } else {
    "[external data directory]"
  }

  if (dir.exists(data_dir)) {
    data_files <- file.path(data_dir, params$data_file)

    if (!file.exists(data_files)) {
      data_notes <- paste(
        "The NACRS data file was not found:",
        file.path(data_dir_display, params$data_file)
      )
    } else {
      # This memo is NACRS-only. Load the selected workbook and expose it as
      # both imported_data_001 and the source-based alias nacrs.
      report_datasets <- tryCatch(
        stats::setNames(
          list(
            read_example_xlsx(data_files)
          ),
          "imported_data_001"
        ),
        error = function(error) {
          data_notes <<- paste(
            "The NACRS example workbook could not be loaded.",
            "Check that it is a valid XLSX file."
          )
          list()
        }
      )

      if (length(report_datasets) == 1L) {
        list2env(report_datasets, envir = knitr::knit_global())
        data_sources <- stats::setNames(data_files, names(report_datasets))

        # Create the source-based alias: synthetic_nacrs.xlsx -> nacrs.
        data_aliases <- local({
          alias_names <- tolower(
            tools::file_path_sans_ext(basename(data_files))
          )
          alias_names <- sub("^(pseudo|synthetic)[._ -]*", "", alias_names)
          alias_names <- make.names(alias_names, unique = TRUE)

          alias_datasets <- stats::setNames(
            unname(report_datasets),
            alias_names
          )
          list2env(alias_datasets, envir = knitr::knit_global())

          stats::setNames(alias_names, names(report_datasets))
        })
      }
    }
  } else {
    data_notes <- paste("Data directory not found:", data_dir_display)
  }
}

```

<style>
.report-meta {
  display: grid;
  grid-template-rows: minmax(0, 1fr);
  grid-auto-flow: column;
  grid-auto-columns: minmax(125px, 1fr);
  gap: 0.45rem;
  align-items: stretch;
  margin: 1.25rem 0 1.5rem;
  padding-bottom: 0.25rem;
  overflow-x: auto;
}
.meta-item {
  min-width: 0;
  padding: 0.7rem 0.9rem;
  border-left: 4px solid #2c7fb8;
  background: #f5f8fa;
  border-radius: 3px;
}
.meta-label {
  display: block;
  color: #657786;
  font-size: 0.78rem;
  text-transform: uppercase;
  letter-spacing: 0.04em;
}
.meta-value {
  display: block;
  overflow-wrap: anywhere;
  font-weight: 600;
}
.report-note {
  margin: 1rem 0 1.5rem;
  padding: 0.8rem 1rem;
  border: 1px solid #f0c36d;
  border-radius: 4px;
  background: #fff8e5;
}
.placeholder {
  color: #6c757d;
  font-style: italic;
}
.table-scroll {
  width: 100%;
  overflow-x: auto;
  margin-bottom: 1.25rem;
}
.summary-table {
  width: 100%;
  margin: 0;
  border-collapse: separate;
  border-spacing: 0;
  font-size: 0.94rem;
}
.summary-table th,
.summary-table td {
  padding: 0.55rem 0.75rem !important;
  border-bottom: 1px solid #e4e9ed;
  vertical-align: top;
  white-space: nowrap;
}
.summary-table thead th {
  position: sticky;
  top: 0;
  z-index: 1;
  color: #29465b;
  background: #eef4f7;
  border-top: 1px solid #d6e0e5;
  border-bottom: 2px solid #9fb8c5;
}
.summary-table tbody tr:nth-child(even) {
  background: #f8fafb;
}
.summary-table tbody tr:hover {
  background: #eef6fa;
}
.continuous-summary-table {
  width: max-content;
  min-width: 100%;
}
.category-table-card {
  max-width: 680px;
  margin: 0.5rem 0 1.75rem;
  overflow: hidden;
  border: 1px solid #d6e0e5;
  border-radius: 8px;
  background: #fff;
  box-shadow: 0 2px 8px rgba(33, 55, 72, 0.06);
}
.category-table-card.long-table {
  max-height: 32rem;
  overflow-y: auto;
}
.category-table-card .summary-table tbody tr:last-child {
  font-weight: 700;
  background: #eaf2f6;
}
.category-table-card .summary-table tbody tr:has(.missing-category) {
  background: #fff8e5 !important;
  box-shadow: inset 4px 0 #e0a12b;
}
.missing-category {
  color: #7a5300;
  font-weight: 700;
}
.missing-na-badge {
  display: inline-block;
  margin-left: 0.4rem;
  padding: 0.08rem 0.35rem;
  border: 1px solid #e2b75e;
  border-radius: 3px;
  background: #ffedbd;
  color: #6c4800;
  font-size: 0.72rem;
  line-height: 1.2;
}
.variable-code {
  margin: -0.35rem 0 0.65rem;
  color: #71818d;
  font-size: 0.86rem;
}
.data-variable-heading {
  margin: 1.25rem 0 0.65rem;
  color: #29465b;
}
.data-disclaimer {
  margin: 0.5rem 0 1.25rem;
  padding: 0.8rem 1rem;
  border-left: 4px solid #2c7fb8;
  border-radius: 4px;
  background: #eef6fa;
  color: #29465b;
}
.data-variable-list {
  display: grid;
  grid-template-columns: repeat(auto-fit, minmax(210px, 1fr));
  gap: 0.45rem 0.75rem;
  margin: 0 0 1.5rem;
  padding: 0;
  counter-reset: data-variable;
  list-style: none;
}
.data-variable-list li {
  display: flex;
  align-items: center;
  min-width: 0;
  padding: 0.45rem 0.6rem;
  counter-increment: data-variable;
  border: 1px solid #e0e7eb;
  border-radius: 5px;
  background: #f8fafb;
}
.data-variable-list li::before {
  content: counter(data-variable);
  flex: 0 0 1.8rem;
  color: #7b8b96;
  font-size: 0.78rem;
}
.data-variable-list code {
  overflow-wrap: anywhere;
  color: #29465b;
  background: transparent;
}
.summary-table caption {
  padding: 0 0 0.65rem;
  color: #71818d;
  text-align: left;
}
</style>

```{r report-metadata, echo=FALSE, results='asis'}
metadata <- c(
  Memo = params$memo,
  Client = params$client,
  `Client Organization` = params$client_organization,
  Version = params$version,
  Status = params$status,
  Classification = params$confidentiality
)
metadata <- metadata[vapply(metadata, is_filled, logical(1))]

if (length(metadata)) {
  items <- paste0(
    '<div class="meta-item"><span class="meta-label">',
    htmltools::htmlEscape(names(metadata)),
    '</span><span class="meta-value">',
    htmltools::htmlEscape(unname(metadata)),
    '</span></div>',
    collapse = "\n"
  )
  cat('<div class="report-meta">', items, '</div>')
}

if (is_filled(params$note)) {
  cat(
    '<div class="report-note"><strong>Note:</strong> ',
    htmltools::htmlEscape(params$note),
    '</div>'
  )
}
```

```{r executive-summary-heading, echo=FALSE, results='asis'}
if (as_flag(params$include_summary)) cat("# Overview\n\n")
```

```{r executive-summary-content, echo=FALSE, results='asis'}
if (as_flag(params$include_summary)) {
  cat("## Introduction\n\n")
  cat(
    paste0(
      '<p>The purpose of this memo is to summarize selected variables from ',
      'the <strong>NACRS</strong> data. It provides a concise description of ',
      'the data source and presents coverage, standardized descriptive ',
      'statistics, and data-quality results.</p>',
      '<p>The memo is organized as follows:</p>',
      '<ul>',
      '<li><strong>Data</strong> identifies the source file and data directory, ',
      'reports the number of records and variables, lists all available ',
      'variables, without displaying row-level records.</li>',
      '<li><strong>Analysis</strong> reports date and fiscal-year coverage, ',
      'summarizes registered measures using type-appropriate methods, and ',
      'presents quality-status distributions and consistency checks.</li>',
      if (as_flag(params$include_appendix)) {
        paste0(
          '<li><strong>Appendix</strong> contains supporting information and ',
          'basic reproducibility details.</li>'
        )
      } else {
        ""
      },
      '</ul>'
    )
  )
}
```

```{r data-overview, echo=FALSE, results='asis'}
if (as_flag(params$include_data_overview)) {
  cat("## Data\n\n")
  cat(
    paste0(
      '<div class="data-disclaimer"><strong>Training data notice:</strong> ',
      'The NACRS data used in this memo are randomly generated synthetic data ',
      'intended solely for training and demonstration purposes. They do not ',
      'represent real patients or actual operational records.</div>'
    )
  )

  if (length(report_datasets) > 0L) {
    overview <- do.call(
      rbind,
      lapply(names(report_datasets), function(dataset_name) {
        dataset <- report_datasets[[dataset_name]]
        is_tabular <- is.data.frame(dataset) || is.matrix(dataset)

        data.frame(
          Dataset = dataset_name,
          Alias = data_aliases[[dataset_name]],
          File = basename(data_sources[[dataset_name]]),
          Rows = if (is_tabular) nrow(dataset) else NA_integer_,
          Columns = if (is_tabular) ncol(dataset) else NA_integer_,
          Class = paste(class(dataset), collapse = ", "),
          check.names = FALSE
        )
      })
    )

    cat("**Data directory:** `", data_dir_display, "`\n\n", sep = "")
    print(knitr::kable(overview, caption = "Loaded NACRS data set"))
    cat("\n")

    data_variable_names <- names(report_datasets[[1L]])
    variable_items <- paste0(
      '<li><code>',
      htmltools::htmlEscape(data_variable_names),
      '</code></li>',
      collapse = ""
    )
    cat(
      '<div class="data-variable-heading"><strong>Variables (',
      length(data_variable_names),
      ')</strong></div>',
      '<ol class="data-variable-list">',
      variable_items,
      '</ol>',
      sep = ""
    )
    cat("\n")

  } else {
    cat(
      '<p class="placeholder">No data sets were loaded from <code>',
      htmltools::htmlEscape(data_dir_display),
      '</code>.</p>',
      sep = ""
    )
  }

  if (length(data_notes) > 0L) {
    cat("### Data-loading notes\n\n")
    cat(
      paste0(
        "- ",
        htmltools::htmlEscape(data_notes),
        collapse = "\n"
      ),
      "\n"
    )
  }
}
```

# Analysis

## NACRS standardized profile

```{r nacrs-standardized-profile, include=FALSE}
nacrs_specification <- load_variable_dictionary(project_dir, "NACRS")
nacrs_profile <- NULL
nacrs_summary_notes <- character()

if (!exists("nacrs", inherits = TRUE) || !is.data.frame(nacrs)) {
  nacrs_summary_notes <- paste(
    "The NACRS profile was not generated because the `nacrs` data set",
    "was not loaded as a data frame."
  )
} else {
  nacrs_profile <- build_dataset_profile(
    nacrs,
    nacrs_specification,
    "NACRS"
  )
}
```

```{r nacrs-standardized-profile-tables, echo=FALSE, results='asis'}
if (!is.null(nacrs_profile)) {
  render_dataset_profile(nacrs_profile, "NACRS")
}

if (length(nacrs_summary_notes) > 0L) {
  cat("## Summary notes\n\n")
  cat(
    paste0(
      "- ",
      htmltools::htmlEscape(nacrs_summary_notes),
      collapse = "\n"
    ),
    "\n"
  )
}
```

```{r appendix, echo=FALSE, results='asis'}
if (as_flag(params$include_appendix)) {
  cat("# Appendix\n\n")
  cat("Add supporting details, methods, definitions, or supplementary results here.\n")
}
```

```{r session-info, echo=FALSE, results='asis'}
if (as_flag(params$include_appendix)) {
  cat("## Reproducibility\n\n")
  cat("Generated with R ", getRversion(), ".\n", sep = "")
}
```
